\textbf{ Spectrophotometric parallaxes with linear models:\\ Accurate distances for luminous red-giant stars }

### 2019-06-02 paper/response_to_referee.txt

During the long period in which we haven't responded to this report,
there was also a paper out by Leung & Bovy. They compare their results
to ours, and we also mention that in the newly added paragraph.

    3 - The authors apply colour cuts to eliminate outliers in the APOGEE
    data, eqs 6 and 7. They also apply a cut in parallax error, eq. 8.,
    along with other quality criteria, to define the training set. So, the
    training set parallax uncertainty distribution is truncated. How do
    these cuts affect the method and its assumptions? In particular, it is
    important to reconcile these selection criteria for the training set
    with the statements presented in the Introduction (page 3):
    "... regressions can be strongly biased if we have a bad selection
    function or data censoring. This means that we cannot apply cuts to
    the Gaia astrometry or signal-to-noise unless those cuts are correctly
    accounted for in our generative model for the data sample. Below, we
    will make no such cuts."  It is in fact clear from Figure 1 that the
    training set is well under represented of low log g stars with respect
    to the parent sample. This already jeopardizes the representativeness
    assumption somewhat.  One useful assessment of the effects of data
    censoring would be to show the spectral S/N or parallax distributions
    of the training data in comparison to that of the entire sample that
    the method is applied to.

This point is important. The Gaia parallax uncertainties are based on
a spacecraft model that is independent of the measurements themselves.
Therefore, cuts on parallax uncertainties are not cuts on parallax and
thus are not subject to the strong censure that we give to such cuts!
But the referee is correct that we have been very unclear about this
and we have attempted to fix this in the relevant paragraph towards the
end of Section 2. This has improved the paper.

The referee is completely correct that the detailed distribution of
stars in parallax, SNR, and log g is not identical in the training set
and the parent sample.  We don't have any reason to believe that this
biases our results. Furthermore, we do have complete coverage. That
is, there are no parallaxes, SNRs, or log g values that exist in the
test set that do not exist in the training set: In each dimension, the
test set is inside the hull of the training set. But in detail there
are small distributional differences. We now comment on this in the
text explicitly in two new sentences in the Data Section. We did not
feel like this merited a figure, but we do have one if the referee
would like to see it.

### 2019-06-02 paper/response_to_referee.txt

To-do items:
------------
- find and kill all HOGG
- find and kill all EILERS

@@

    The paper presents a method to estimate spectrophotometric parallaxes
    based only on input data, without adoption of physically motivated
    models and priors. The paper is written in an way that is strongly
    focused on the method itself, but leaves several complementary aspects
    that are certainly important to assess the reliability of the inferred
    parallaxes and, most importantly, distances. I recommend its
    publication only after such major issues are addressed.

    Major issues

    1 - The data-driven method described in the paper details out how
    spectrophotometric parallaxes are inferred. But the paper does not
    give the reader a clue about how spectrophotometric distances are
    estimated. The authors make it clear in Section 2 that they never
    invert *observed* parallaxes to estimate distances in their
    method. But are the spectrophotometric distances they infer (shown in
    Figures 5 and 6) just the inverse of the *inferred* parallaxes? If so,
    this should be made explicit in the text, along with some discussion
    about the problems (or lack of) that may arise from this
    inversion. Even though the spectrophotometric parallaxes themselves
    are always positive, a reader would like to be convinced that the
    distances shown in Figures 5 and 6 do not suffer from biasses caused
    by the asymmetric nature of the distances distribution that results
    from a symmetric error distribution in parallaxes (as stated by the
    likelihood assumption of the method). How do these biasses simply go
    away when the method's w(sp) +/- sigma(sp) are used? In particular, a
    negative parallax for many stars may be consistent with their w(sp)
    +/- sigma(sp).

These issues are subtle! We are assuming that the GAIA-MEASURED
PARALLAXES have noise that is symmetric in parallax space. In our
data-driven model, we use a linear combination of features to predict
or estimate a LOG PARALLAX, so if anything we are claiming that our
estimates (not the data, but our estimates based on the data) are
symmetric in log parallax, not parallax. But we aren't even really
claiming that. The issues are subtle. However, this does deserve some
discussion and we have added that to the captions of Figure 5 and 6,
and in the last paragraph of the discussion, for extra-special
emphasis.  The short answer is that we just exponentiated the negative
of our log-parallax estimates to make those figures.

### 2019-06-03 paper/sp.tex

The Training Sets are shown in the bottom panels of \figurename~\ref{fig:samples}.
Comparison of the top and bottom panels of {\figurename}~\ref{fig:samples} shows
that there is more color range in the Parent Sample than in the Training Set.
This challenges the representativeness assumption in \sectionname~\ref{sec:assumptions}.
The principal reason for the difference is that the \gaia\ quality cuts
exclude stars preferentially from crowded regions, which also tend to be the
most dust-obscured.
The hope of the model assumptions is that the Training Set will contain sufficient
dust variation that the model will naturally learn the dust corrections and extrapolate
acceptably.

>>>>

The Training Sets are shown in the bottom panels of \figurename~\ref{fig:samples}.
Comparison of the top and bottom panels of {\figurename}~\ref{fig:samples} shows
that there is more color range in the Parent Sample than in the Training Set.
This challenges the representativeness assumption in \sectionname~\ref{sec:assumptions}.
The principal reason for the difference is that the \gaia\ quality cuts
exclude stars preferentially from crowded regions, which also tend to be the
most dust-obscured.
The hope of the model assumptions is that the Training Set will contain sufficient
dust variation that the model will naturally learn the dust corrections and extrapolate
acceptably.
In terms of the more mundane properties of \gaia\ parallax $\gparallax$,
\apogee\ spectroscopic signal-to-noise ratio,
or surface gravity $\logg$, there are small differences between the distributions
in the Training Sets and the Parent Sample.
However, in each of these dimensions, the Training Set do span the full range
of the Parent Sample, so the Training Set does have full coverage in the
space.
